Adaptive Waveform Inversion for Transmitted Wavefields

### 2024-10-29 repro/papers/awi/paper.tex
near final

Various modifications of straighforward best-fit FWI have been
suggested to overcome this ``cycle-skipping'' pathology. One of these
is Adaptive Waveform Inversion \cite[]{Warner:16} (AWI). This
technique has been successfully applied to industry-scale seismic data
\cite[]{GuaschWarnerRavaut:GEO19,Warneretal:SEG21} and shows promise for
ultrasonic imaging of the the human brain
\cite[]{Guaschetal:NPJDM20}.

@@

The reader will
note that a normalization component of the AWI objective function
definition is important in reaching this conclusion. The connection
between objective normalization and the asymptotic equivalence to
travel time inversion for transmission data has nothing to do with the
reasons for normalizing the objective previoiusly advanced in the
literature \cite[]{Warner:16,LiAlkhalifah:21}. Another modification of
FWI based on similar ideas, Matched Source Waveform Inversion (MSWI)
(\cite{HuangSymes2015SEG,HuangSymes:Geo17}) lacks the AWI
normalization, and has a looser relation to travel time inversion. The
arguments presented here for AWI (and MSWI) are refinements of those
given by \cite{HuangSymes:Geo17}. which were in turn based on ideas
presented in Song's thesis \cite[]{Song:94c}.

The conclusions of this study depend on a number of limitations on the
acoustic model of wave propagation and the formulation of the related
inverse problem. The principal constraints are:
\begin{itemize}
  \item energy sources are modeled as isotropic point radiators, with
    known time dependence vanishing outside a finite time window;
  \item mechanical parameter fields (coefficients in the acoustic wave
    equation) are smooth, that is, infinitely differentiable as
    functions of position, and the material density is known and
    position-independent: only the bulk modulus (or wave velocity) is
    to be determined;
  \item waves propagate throughout Euclidean space-time - that is, no
    physical boundaries are modeled;
  \item the data of the inverse problem are the restrictions of the
    pressure field to a finite number of points in space (receiver
    positions), for a finite number of source positions and all time;
  \item local energy decays exponentially - that is, the pressure and
    all of its derivatives tend to zero at an exponential rate as $t
    \rightarrow \infty$,
    uniformly in any bounded spatial domain.
  \end{itemize}
  These constraints are discussed in detail in the next section. Some
  could certainly be relaxed. For instance, exponential decay of local
  energy implies that the finite time duration of actual measured data
  can be treated as a perturbation of the infinite time data setting
  posed here. However these hypotheses enable a simple and clear
  analysis of AWI, and seem a reasonable starting point for further
  investigation.

@@

The parameter to be determined in the inverse problem studied here is
the bulk modulus $\kappa$. It is presumed is bounded and smooth:
\begin{equation}
  \label{eqn:coeff}
  \log \kappa \in C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3)
\end{equation}
Under these hypotheses, the symmetric hyperbolic system \ref{eqn:awe}
has a unique distribution
solution if $f \in \cal{E}'(\bR)$ (for example \cite{Lax:PDENotes},
Theorem 7.4). The Propagation of
Singularities Theorem (\cite{Tay:81}, Theorem VI.2.1) and
the pull-back property of distributions \cite{Dui:95}, Proposition 1.3.3)
imply that if $f \in \cal{E}'(\bR)$ and $(\bx_s,\bx_r) \in X_{sr}$,
then the trace $p(\bx_r, \cdot;\bx_s)$ on $\{(\bx_r,t): t \in \bR\}$ is well-defined,
as a distribution vanishing for sufficiently large
negative $t$. Define the {\em modeling operator}
\begin{equation}
  \label{eqn:modop}
  F: C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3)
  \rightarrow X_{sr} \times {\cal D}'(\bR)
\end{equation}
by
\begin{equation}
  \label{eqn:modopdef}
  F[\kappa]= p|_{X_{sr} \times \bR}
\end{equation}

@@

Under conditions G1 - G4, the construction described by
\cite{Friedlander:75} produces for each $m \in M$ functions $\tau[\kappa]$ (travel time), $a[\kappa]$
(geometric amplitude) on $X_s \times \Omega$ and $b[\kappa]$
(remainder kernel) on $X_s \times \Omega \times \bR$, satisfying
\begin{itemize}
\item[H1. ] $\tau[\kappa], a[\kappa],$ and $b[\kappa]$ are smooth off the diagonal,
that is, $\tau[\kappa](\bx_s,\cdot) \in C^{\infty}(\Omega \setminus \{\bx_s\})$
and similarly for $a[\kappa],b[\kappa]$;
\item[H2. ] there exists $A>0$ so that for any $\log \kappa \in M$
  and $(\bx_s,\bx_r) \in X_{sr}$,
  $|\log a[\kappa](\bx_s,\bx_r)| < A$;
\item[H3. ] for any $k \ge 0$, there exists $B \ge 0,\delta > 0$ so
  that for any $(\bx_s,\bx_r) \in X_{sr}$, $\kappa \in M$, 
\begin{equation}
  |\partial^k_t b[\kappa](\bx_r,t;\bx_r,s)| \le B e^{-\delta t}
  \label{eqn:best}
\end{equation}
\end{itemize}
(The exponential decay estimate in item H3 follows from hypothesis G4.)

@@

@@

AWI is based an {\em extension} of the modeling operator $F$, that is,
an operator ${\bar F}$ with larger domain, coinciding with $F$ on its domain..
The extended modeling operator ${\bar F}$ maps extended models
$\bar{m}=(\kappa,\bar{w})$ incorporating extended sources
$\bar{w}({\bf x}_r,t;{\bf x}_s)$ to the same sampling of the pressure
field. Denote by $\bar{p}(\bx,t;\bx_s,\bx_r), \bar{\bv}(\bx,t;\bx_s,\bx_r)$
the solution of the acoustic system \ref{eqn:awe} with $w(t)$ in the
right-hand side of the first equation replaced by
$\bar{w}(\bx_r,\cdot;\bx_s)$. Then
$\bar{F}[\kappa,\bar{w}](\bx_r,t;\bx_s) =
\bar{p}(\bx_r,t;\bx_s,\bx_r)$. As was the case for $F$, the Hadamard
decomposition \ref{eqn:green} and conditions G1-G4, W1-W3 imply that
$\bar{F}[\kappa, \cdot]$ is a bounded map: $L^2(X_{sr} \times \bR)
\rightarrow L^2(X_{sr} \times \bR)$. If all  
of the wavelets are the same, that is, ${\bar w}({\bf  
  x}_r,t;{\bf x}_s) = w(t)$ is independent of source and receiver  
position, then ${\bar F}[\kappa,\bar{w}] = F[\kappa]$. That is, $F$ is a special
case, or restriction, of ${\bar F}$, so ${\bar F}$ is an extension of
$F$. This is the {\em source-receiver} extension, in the terminology of \cite{HuangSymes2015SEG}.

\cite{Warner:16} represent the extended sources $\bar{w}$ as time convolutions of the
(known) source $w$ with an {\em adaptive filter} $u \in L^2(X_{sr}
\times \bR)$, that is, $\bar{w}=u * w$.
Here, $u*w$ denotes the result of convolving each trace in $u$ by $w
\in L^2(\bR)$. I
will also use the asterisk to denote the trace-by-trace convolution
of two data sets with the same set of source and receiver coordinates,
that is, the data set consisting of time convolutions of traces with the same source
and receiver coordinates. The meaning will be clear from context.

For such extended sources,
\[
  \bar{F}[\kappa,u*w] = u*F[\kappa]
\]
Introduce the notation $S[\kappa]$ for the operator of convolution
with $F[\kappa]$, that is,
\begin{equation}
  \label{eqn:sdef}
  S[\kappa]u = u*F[\kappa]
\end{equation}
Then $S[\kappa]$ is bounded on $L^2(X_{sr} \times \bR)$, uniformly in
$\log \kappa \in M$, thanks to
the bound \ref{eqn:fwdnorm} for $p=1$. $S$ also represents an
extension of $F$, in the sense that with the choice
$u(\bx_r,t;\bx_s)=\delta(t)$, the output of $F$ is recovered:
\begin{equation}
  \label{eqn:sconsist}
  S[\kappa]\delta = F[\kappa].
\end{equation}

@@

\cite{Warner:16} assume (implicitly) that there is always a unique
solution of the extended inversion problem, formulated in terms of
$S$: that is, that
for any $\kappa$ and any data $d$, there is an
adaptive filter $u[\kappa;d]$ for which
$S[\kappa]u[\kappa;d]= d$. If $d$ is noise-free, that is, $d=F[\kappa]$, then the consistency relation \ref{eqn:sconsist}
suggests that $u[\kappa;d](\bx_r,t;\bx_s) = \delta(t)$. This ideal adaptive filter is in the null
space of multiplication by $t$, that is, the operator
$u(\bx_r,t;\bx_s) \mapsto t u(\bx_r,t;\bx_s)$. This
observation motivates the definition of the AWI objective function
(\cite{Warner:16}, equation XXXX)::
\begin{equation}
  \label{eqn:awideforig}
  J_{\rm AWI}[m;d] = \sum_{\bx_s,\bx_r} \left(\frac{\int dt |tu[\kappa;d](\bx_r,t;\bx_s)|^2}{\int dt |u[\kappa;d](\bx_r,t;\bx_s)|^2}\right) 
\end{equation}
in which $u[\kappa;d]$ is as defined above, that is, the solution of $S[\kappa]u[\kappa;d]
= d$. \cite{Warner:16} also formulate
AWI in terms of the inverse filter, that is, $u$ for which
$u*d = F[\kappa]$, and in fact recommend that definition. The present
discussion uses only the forward filter.
Adaptive waveform inversion consists in iterative minimization of
$J_{\rm AWI}$: choose an initial estimate $\kappa_0$ of the
model, then update $\kappa$ repeatedly using a gradient
descent method. \cite{Warner:16} describes the computation of $J_{\rm AWI}$
and its gradient in detail, and their use in a quasi-Newton
optimization algorithm.

The same observation motivates the definition of the {\em Matched
  Source Waveform Inversion} (MSWI) objective studied by
\cite{HuangSymes2015SEG,HuangSymes:Geo17}:
\begin{equation}
  \label{eqn:mswideforig}
  J_{\rm MSWI}[m;d] = \sum_{\bx_s,\bx_r} \int dt |tu[\kappa;d](\bx_r,t;\bx_s)|^2
\end{equation}
That is, the two objectives differ only by the trace-by-trace
normalization featured in the definition of AWI (equation
\ref{eqn:awideforig}). Much of the analysis of $J_{\rm AWI}$ pertains to
$J_{\rm MSWI}$, as will be explained below.
The trace-by-trace normalization in the definition of $J_{\rm AWI}$
will prove quite consequential.

@@

In fact, these two defects are really symptoms of the instability of
deconvolution for kernels with the properties of $F[\kappa]$.  To
begin with, $S[\kappa] = F[\kappa]*$ defines a bounded map on
$L^2(X_{sr}\times \bR)$, since from \ref{eqn:fwdgreen} and the
definition \ref{eqn:sdef} of  $S[\kappa]$,
\[
  S[\kappa]u = \rho( a[\kappa]\delta    +  (b[\kappa]H)) *\delta_{\tau[\kappa]}*\partial_t f *u,
\]
\begin{equation}
  \label{eqn:sgood}
  = \rho(a[\kappa]\delta_{\tau[\kappa]}*\partial_t f +
  (b[\kappa](0)\delta + (\partial_t b[\kappa])H)
  *\delta_{\tau[\kappa]}*f) * u
\end{equation}
The key properties of $S$ follow directly from Young's inequality and
the representation \ref{eqn:fwdgreen}:
\begin{lemma}
  \label{thm:snorm}
  Under the assumptions leading to \ref{eqn:fwdgreen} and
  the properties F1 and F2, $S[\kappa]$ is a bounded operator on
  $L^2(X_{sr} \times \bR)$, and there exists $C>0$ so
  that for any $\kappa \in M$,
  \begin{equation}
    \label{eqn:snorm}
    \|S[\kappa]\| \le C\|\partial_t f\|_{L^1(\bR)}.
  \end{equation}
\end{lemma}

The conventional remedy for instability in a deconvolution problem of
this type is so-called ``pre-whitening'', as
mentioned by \cite{Warner:16} ``Pre-whitening'' is a version of
Tihonov regularization. It  amounts to replacement of the equation
$S[\kappa]u=d$ by the regularized least squares problem: 
\begin{equation}
  \label{eqn:reg}
  u_{\sigma}[\kappa;d] = \mbox{argmin}_{u \in L^2(X_{sr} \times \bR)} (\|S[\kappa]u-d\|^2 + \sigma\|u\|^2)
\end{equation}

@@

This definition has introduced a new defect: while the denominator is
well-defined for $\log \kappa \in M$, $d \times L^2(X_{sr} \times
\bR)$ for $\sigma > 0$, the numerator may not be, since there is no
guarantee that $Tu_{\sigma}[\kappa;d] \in L^2(X_{sr} \times
\bR)$.

@@

  Thus
  \[
    S^TS =
    \rho^2(a\delta  +  (\check{b}\check{H})) *\delta_{-\tau}*\partial_t
    f * (a\delta  +  (bH)) *\delta_{\tau}*\partial_t f*
  \]
  \[
    = \rho^2((a\delta  +  (\check{b}\check{H}))*(a\delta  +
    (bH))*\partial_t f *\partial_t f*
  \]
  \[
    [T,S^TS] = \rho^2((T\check{b}\check{H}) *(a\delta  +
    (bH))*+ (a\delta  +  (\check{b}\check{H}))*(TbH))*\partial_t f
    *\partial_t f*
  \]
  \[
    +2\rho^2((a\delta  +  (\check{b}\check{H}))*(a\delta  +
    (bH)))* \partial_t f  *T\partial_t f*
  \]
  \[
    =\rho^2((T\partial_t\check{b}\check{H}) *(a\delta  +
    (bH))*+ (a\delta  +  (\check{b}\check{H}))*(T\partial bH))* f
    *\partial_t f*
   \]
  \begin{equation}
    \label{eqn:tsscomm}
     +2\rho^2((a\delta  +  (\check{b}\check{H}))*(a\delta  +
    (bH)))* \partial_t f  *T\partial_t f*.
  \end{equation}

  Referring to the exponential decay assumption \ref{eqn:best}
  and properties F1, F2 of $f$, 
  observe that all summands on the RHS of identity \ref{eqn:tsscomm} are in
  $L^1(X_{sr} \times \bR)$, so $[T,S^TS]$ restricts to a bounded
  operator on $L^2(X_{sr}\times \bR)$, with
  \begin{equation}
    \label{eqn:tsscommnorm}
    \|[T,S^TS]\| \le C\|\partial_t f\|_{L^1(\bR)}(\|f\|_{L^1(\bR)}+\|T\partial_tf\|_{L^1(\bR)})
  \end{equation}

@@

  Multiplying the normal equation for $u_{\sigma}$Rearranging,
  \[
    (S^TS+\sigma I)Tu_{\sigma} = -[T,S^TS]u_{\sigma} + [T,S^T]d + S^TTd.
  \]
  A theorem of von Neumann (\cite{Yosida}, VII.3, Theorem 2) implies
  that $S^TS+\sigma I$ has a bounded inverse, so the distribution
  $Tu_{\sigma} \in L^2(X_{sr} \times \bR)$. The inverse is bounded by
  $1/\sigma$, so the estimates \ref{eqn:snorm}, \ref{eqn:unorm},
  \ref{eqn:tcommst}, and \ref{eqn:tsscommnorm} combine to yield the
  bound \ref{eqn:tunorm}.

@@

In particular, the modified AWI objective $J_{\rm AWI,\lambda}$ is {\em exactly} an approximation to
the travel-time tomography objective, with errors proportional to the squared
wavelength proxy $\lambda$.

@@

\begin{prop} Under the hypotheses of Proposition \ref{thm:leading},
  \begin{eqnarray}
    \label{eqn:diffests}
    \|\ul-\uzl\| &\le& C \nonumber \\
    \|T(\ul-\uzl)\| &\le& C
  \end{eqnarray}
\end{prop}

>>>>

Various modifications of straighforward best-fit FWI have been
suggested to overcome this ``cycle-skipping'' pathology. Many of these
modifications employ {\em model extension}, that is, the introduction
of a larger space of model parameters than derives from basic
physics. An example is the {\em source-receiver} extension \cite[]{HuangSymes:17}, in which
source waveforms are permitted to depend (unphysically) on receiver
locations. The focus of this paper is Adaptive Waveform Inversion
\cite[]{Warner:16} (AWI), a modification of FWI using the
source-receiver extension. AWI has been successfully applied to industry-scale seismic data
\cite[]{GuaschWarnerRavaut:GEO19,Warneretal:SEG21} and shows promise for
ultrasonic imaging of the the human brain
\cite[]{Guaschetal:NPJDM20}.

@@

The reader will
note that a normalization component of the AWI objective function
definition is important in reaching this conclusion. The connection
between objective normalization and the asymptotic equivalence to
travel time inversion for transmission data has nothing to do with the
reasons for normalizing the objective previoiusly advanced in the
literature \cite[]{Warner:16,LiAlkhalifah:21}. Another modification of
FWI based on source-receiver extension, Matched Source Waveform Inversion (MSWI)
\cite[]{HuangSymes2015SEG,HuangSymes:Geo17} lacks the AWI
normalization, and has a looser relation to travel time inversion. The
arguments presented here for AWI (and MSWI) are refinements of those
given by \cite{HuangSymes:Geo17}. which were in turn based on ideas
presented in Song's thesis \cite[]{Song:94c}.

The conclusions of this study depend on a number of limitations on the
acoustic model of wave propagation and the formulation of the related
inverse problem. The principal constraints are:
\begin{itemize}
\item energy sources are modeled as isotropic point radiators, with
  known time dependence vanishing outside a finite time window;
\item mechanical parameter fields (coefficients in the acoustic wave
  equation) are smooth, that is, infinitely differentiable as
  functions of position, and the material density is known and
  position-independent: only the bulk modulus (or wave velocity) is
  to be determined;
\item waves propagate throughout Euclidean space-time - that is, no
  physical boundaries are modeled;
\item the data of the inverse problem are the restrictions of the
  pressure field to a finite number of points in space (receiver
  positions), for a finite number of source positions and all time;
\item local energy decays exponentially - that is, the pressure and
  all of its derivatives tend to zero at an exponential rate as $t
  \rightarrow \infty$,
  uniformly in any bounded spatial domain.
\item only one ray of geometric
  acoustics connects each source-receiver location pair (single arrivals)
\end{itemize}

These constraints are discussed in detail in the next section. Some
could certainly be relaxed. For instance, exponential decay of local
energy implies that the finite time duration of actual measured data
can be treated as a perturbation of the infinite time data setting
posed here. The exception is the single-arrival hypothesis: it is
essential, as without it, the link between the travel time and
extended waveform inversion is lost. This is a feature of the
source-receiver extension, not of a particular objective function
(MSWI, AWI,...) formulated in terms of it. This extension treats each receiver signal
independently. If only a single identifiable waveform arrives at
each receiver, AWI (and similar approaches) can implicitly identify
its arrival time, and the link to travel time inversion is
established. If distinct waveforms are present at multiple times,
the single receiver signal does not determine which waveform in the
predicted signal should align with which waveform in the data to be
fitted. This ``crosstalk'' ambiguity is analyzed in the fifth
section, and afflicts all approaches based on the source-receivre
extension.

In any case, these hypotheses enable a simple and clear
analysis of AWI, and seem a reasonable starting point for further
investigation.

@@

The parameter to be determined in the inverse problem studied here is
the bulk modulus $\kappa$. It is presumed is bounded and smooth:
\begin{equation}
  \label{eqn:coeff}
  \log \kappa \in C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3)
\end{equation}
Under these hypotheses, the symmetric hyperbolic system \ref{eqn:awe}
has a unique distribution
solution if $f \in \cal{E}'(\bR)$ (for example \cite{Lax:PDENotes},
Theorem 7.4). The Propagation of
Singularities Theorem (\cite{Tay:81}, Theorem VI.2.1) and
the pull-back property of distributions (\cite{Dui:95}, Proposition 1.3.3)
imply that if $f \in \cal{E}'(\bR)$ and $(\bx_s,\bx_r) \in X_{sr}$,
then the trace $p(\bx_r, \cdot;\bx_s)$ on $\{(\bx_r,t): t \in \bR\}$ is well-defined,
as a distribution vanishing for sufficiently large
negative $t$. Define the {\em modeling operator}
\begin{equation}
  \label{eqn:modop}
  F: C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3)
  \rightarrow X_{sr} \times {\cal D}'(\bR)
\end{equation}
by
\begin{equation}
  \label{eqn:modopdef}
  F[\kappa]= p|_{X_{sr} \times \bR}
\end{equation}

@@

Under conditions G1 - G4, the construction described by
\cite{Friedlander:75} produces for each $m \in M$ functions $\tau[\kappa]$ (travel time), $a[\kappa]$
(geometric amplitude) on $X_s \times \Omega$ and $b[\kappa]$
(remainder kernel) on $X_s \times \Omega \times \bR$, satisfying
\begin{itemize}
\item[H1. ] $\tau[\kappa], a[\kappa],$ and $b[\kappa]$ are smooth off the diagonal,
  that is,
  (i) for each $\bx_s \in X_s$, $\tau[\kappa](\bx_s,\cdot) \in C^{\infty}(\Omega \setminus
  \{\bx_s\})$, and similarly for $a[\kappa],b[\kappa]$; $\tau$
  satisfies the eikonal equation
  \[
    \sqrt{\frac{\kappa}{\rho}} |\nabla \tau(\bx_s,\cdot)| = 1 \mbox{ in } \Omega \setminus
    \{\bx_s\}, \, \lim_{\bx \rightarrow \bx_s} \tau(\bx_s,\bx) = 0.
  \]
\item[H2. ] there exists $A>0$ so that for any $\log \kappa \in M$
  and $(\bx_s,\bx_r) \in X_{sr}$,
  $|\log a[\kappa](\bx_s,\bx_r)| < A$;
\item[H3. ] for any $k \ge 0$, there exists $B \ge 0,\delta > 0$ so
  that for any $(\bx_s,\bx_r) \in X_{sr}$, $\kappa \in M$, 
\begin{equation}
  |\partial^k_t b[\kappa](\bx_r,t;\bx_r,s)| \le B e^{-\delta t}
  \label{eqn:best}
\end{equation}
\end{itemize}
(The exponential decay estimate in item H3 follows from hypothesis G4.)

@@

Generally, constants denoted $C$ etc. will absorb any dimensional
scale factors necessary to make sense of the expressions in which they
appear. I will call out those few cases in which explicit use of
dimensinoal scale factors is necessary to avoid irretrievable
dimensional confusion.

@@

AWI is based an {\em extension} of the modeling operator $F$, that is,
an operator ${\bar F}$ with larger domain, coinciding with $F$ on its domain..
The extended modeling operator ${\bar F}$ maps extended models
$\bar{m}=(\kappa,\bar{f})$ incorporating extended sources
$\bar{f}({\bf x}_r,t;{\bf x}_s)$ to the same sampling of the pressure
field. Denote by $\bar{p}(\bx,t;\bx_s,\bx_r), \bar{\bv}(\bx,t;\bx_s,\bx_r)$
the solution of the acoustic system \ref{eqn:awe} with $f(t)$ in the
right-hand side of the first equation replaced by
$\bar{f}(\bx_r,\cdot;\bx_s)$. Then
$\bar{F}[\kappa,\bar{f}](\bx_r,t;\bx_s) =
\bar{p}(\bx_r,t;\bx_s,\bx_r)$. As was the case for $F$, the Hadamard
decomposition \ref{eqn:green} and conditions G1-G4, W1-W3 imply that
$\bar{F}[\kappa, \cdot]$ is a bounded map: $L^2(X_{sr} \times \bR)
\rightarrow L^2(X_{sr} \times \bR)$. If all  
of the wavelets are the same, that is, ${\bar f}({\bf  
  x}_r,t;{\bf x}_s) = f(t)$ is independent of source and receiver  
position, then ${\bar F}[\kappa,\bar{f}] = F[\kappa]$. That is, $F$ is a special
case, or restriction, of ${\bar F}$, so ${\bar F}$ is an extension of
$F$. This is the {\em source-receiver} extension, in the terminology of \cite{HuangSymes2015SEG}.

\cite{Warner:16} represent the extended sources $\bar{f}$ as time convolutions of the
(known) source $f$ with an {\em adaptive filter} $u \in L^2(X_{sr}
\times \bR)$, that is, $\bar{f}=u * f$.
Here, $u*f$ denotes the result of convolving each trace in $u$ by $f$. I
will also use the asterisk to denote the trace-by-trace convolution
of two data sets with the same set of source and receiver coordinates,
that is, the data set consisting of time convolutions of traces with the same source
and receiver coordinates. The meaning will be clear from context.

For such extended sources,
\[
  \bar{F}[\kappa,u*f] = u*F[\kappa]
\]
Introduce the notation $S[\kappa]$ for the operator of convolution
with $F[\kappa]$, that is,
\begin{equation}
  \label{eqn:sdef}
  S[\kappa]u = u*F[\kappa]
\end{equation}
$S$ represents an
extension of $F$, in the sense that with the choice
$u(\bx_r,t;\bx_s)=\delta(t)$, the output of $F$ is recovered:
\begin{equation}
  \label{eqn:sconsist}
  S[\kappa]\delta = F[\kappa].
\end{equation}

@@

\cite{Warner:16} assume (implicitly) that there is always a unique
solution of the extended inversion problem, formulated in terms of
$S$: that is, that
for any $\kappa$ and any data $d$, there is an
adaptive filter $u[\kappa;d]$ for which
$S[\kappa]u[\kappa;d]= d$. If $d$ is noise-free, that is, $d=F[\kappa]$, then the consistency relation \ref{eqn:sconsist}
suggests that $u[\kappa;d](\bx_r,t;\bx_s) = \delta(t)$. This ideal adaptive filter is in the null
space of multiplication by $t$, that is, the operator
$u(\bx_r,t;\bx_s) \mapsto t u(\bx_r,t;\bx_s)$. This
observation motivates the definition of the AWI objective function
(\cite{Warner:16}, equation 4)::
\begin{equation}
  \label{eqn:awideforig}
  J_{\rm AWI}[\kappa;d] = \sum_{\bx_s,\bx_r} \left(\frac{\int dt |tu[\kappa;d](\bx_r,t;\bx_s)|^2}{\int dt |u[\kappa;d](\bx_r,t;\bx_s)|^2}\right) 
\end{equation}
in which $u[\kappa;d]$ is as defined above, that is, the solution of $S[\kappa]u[\kappa;d]
= d$. \cite{Warner:16} also formulate
AWI in terms of the inverse filter, that is, $u$ for which
$u*d = F[\kappa]$, and in fact recommend that definition. The present
discussion uses only the forward filter.
Adaptive waveform inversion consists in iterative minimization of
$J_{\rm AWI}$: choose an initial estimate $\kappa_0$ of the
model, then update $\kappa$ repeatedly using a gradient
descent method. \cite{Warner:16} describes the computation of $J_{\rm AWI}$
and its gradient in detail, and their use in a quasi-Newton
optimization algorithm.

The same observation motivates the definition of the {\em Matched
  Source Waveform Inversion} (MSWI) objective studied by
\cite{HuangSymes2015SEG,HuangSymes:Geo17}:
\begin{equation}
  \label{eqn:mswideforig}
  J_{\rm MSWI}[\kappa;d] = \sum_{\bx_s,\bx_r} \int dt |tu[\kappa;d](\bx_r,t;\bx_s)|^2
\end{equation}
That is, the two objectives differ only by the trace-by-trace
normalization featured in the definition of AWI (equation
\ref{eqn:awideforig}). Much of the analysis of $J_{\rm AWI}$ pertains to
$J_{\rm MSWI}$, as will be explained below.
The trace-by-trace normalization in the definition of $J_{\rm AWI}$
will prove quite consequential.

@@

In fact, these two defects are really symptoms of the instability of
deconvolution for kernels with the properties of $F[\kappa]$.  From \ref{eqn:fwdgreen} and the
definition \ref{eqn:sdef} of  $S[\kappa]$,
\[
  S[\kappa]u = \rho( a[\kappa]\delta    +  (b[\kappa]H)) *\delta_{\tau[\kappa]}*\partial_t f *u,
\]
\begin{equation}
  \label{eqn:sgood}
  = \rho(a[\kappa]\delta_{\tau[\kappa]}*\partial_t f +
  (b[\kappa](0)\delta + (\partial_t b[\kappa])H)
  *\delta_{\tau[\kappa]}*f) * u.
\end{equation}
Since $\partial_t f$ is even, the formal adjoint $S[\kappa]^T$ is given by
\begin{equation}
  \label{eqn:sadjgood}
  S[\kappa]^Tu = \rho( a[\kappa]\delta    +  (\check{b}[\kappa]\check{H}))
  *\delta_{-\tau[\kappa]}*\partial_t f *u,
\end{equation}
in which $\check{u}(t) = u(-t)$.
The key properties of $S$ follow directly from Young's inequality and
the representation \ref{eqn:sgood}:
\begin{lemma}
  \label{thm:snorm}
  Under the assumptions G1-G4 leading to \ref{eqn:fwdgreen} and
  the properties F1 and F2, the expressions \ref{eqn:sgood} and \ref{eqn:sadjgood} define 
  bounded operators on
  $\Pi_{X_{sr}} {\cal S}(\bR)$ and on
  $L^2(X_{sr} \times \bR)$, and there exists $C>0$ so
  that for any $\kappa \in M$,
  \begin{equation}
    \label{eqn:snorm}
    \|S[\kappa]\| \le C\|\partial_t f\|_{L^1(\bR)}.
  \end{equation}
\end{lemma}
\begin{proof} From the estimates H3, $b[\kappa](\bx_r,\cdot;\bx_s)H *
  \partial_tf \in {\cal S}(\bR)$ for all $(\bx_s,\bx_r) \in X_{sr}$,
  whence the first conclusion. Bounded extension to $L^2(X_{sr} \times
  \bR)$ follows from Young's inequality.
\end{proof}

The conventional remedy for instability in a deconvolution problem of
this type is so-called ``pre-whitening'', as
mentioned by \cite{Warner:16} ``Pre-whitening'' is a version of
Tihonov regularization. It  amounts to replacement of the equation
$S[\kappa]u=d$ by the regularized least squares problem: 
\begin{equation}
  \label{eqn:reg}
  u_{\sigma}[\kappa;d] = \mbox{argmin}_{u \in L^2(X_{sr} \times \bR)} (\|S[\kappa]u-d\|^2 + \sigma\|u\|^2)
\end{equation}
For future reference, examine the
implications of definition \ref{eqn:reg} for the units of the regularization weight $\sigma$,
which  is
necessarily dimensional. Since the data $d$ has dimensions of
pressure, so must $S[m]u = F[m]*u$. Of course $F[m]$ is also sampled
pressure, and convolution is a $t$ integration, so for $S[m]u$ to be a
pressure, $u$ must have dimensions of 1/time. Therefore
$\|S[m]u-d\|^2$ has dimensions of pressure$^2 \times $time, $\|u\|^2$ has
dimensions of 1/time, whence $\sigma$ must be dimensioned as
pressure$^2 \times$  time$^2$.

@@

The definitions \ref{eqn:awidefbis} and \ref{eqn:mswidef} have introduced a new defect: there is no
guarantee that $Tu_{\sigma}[\kappa;d] \in L^2(X_{sr} \times\bR)$.

@@

  Thus
  \[
    S^TS =
    \rho^2(a\delta  +  (\check{b}\check{H})) *\delta_{-\tau}*\partial_t
    f * (a\delta  +  (bH)) *\delta_{\tau}*\partial_t f*
  \]
  \[
    = \rho^2((a\delta  +  (\check{b}\check{H}))*(a\delta  +
    (bH))*\partial_t f *\partial_t f*.
  \]
  $T$ is also a bounded map on $\Pi_{X_{sr}} {\cal S}(\bR)$, as is
  its commutator with $S^TS$, and 
  \[
    [T,S^TS] = \rho^2((T\check{b}\check{H}) *(a\delta  +
    (bH))*+ (a\delta  +  (\check{b}\check{H}))*(TbH))*\partial_t f
    *\partial_t f*
  \]
  \[
    +2\rho^2((a\delta  +  (\check{b}\check{H}))*(a\delta  +
    (bH)))* \partial_t f  *T\partial_t f*
  \]
  \[
    =\rho^2((T\partial_t\check{b}\check{H}) *(a\delta  +
    (bH))*+ (a\delta  +  (\check{b}\check{H}))*(T\partial bH))* f
    *\partial_t f*
   \]
  \begin{equation}
    \label{eqn:tsscomm}
     +2\rho^2((a\delta  +  (\check{b}\check{H}))*(a\delta  +
    (bH)))* \partial_t f  *T\partial_t f*.
  \end{equation}

  Referring to the exponential decay assumption \ref{eqn:best}
  and properties F1, F2 of $f$, 
  observe that all summands on the RHS of identity \ref{eqn:tsscomm} are in
  $L^1(X_{sr} \times \bR)$, so $[T,S^TS]$ extends to a bounded
  operator on $L^2(X_{sr}\times \bR)$, with
  \begin{equation}
    \label{eqn:tsscommnorm}
    \|[T,S^TS]\| \le C\|\partial_t f\|_{L^1(\bR)}(\|f\|_{L^1(\bR)}+\|T\partial_tf\|_{L^1(\bR)})
  \end{equation}

@@

  Multiplying the normal equation for $u_{\sigma}$ by $T$ and earranging,
  \[
    (S^TS+\sigma I)Tu_{\sigma} = -[T,S^TS]u_{\sigma} + [T,S^T]d + S^TTd.
  \]
  A theorem of von Neumann (\cite{Yosida}, VII.3, Theorem 2) implies
  that $S^TS+\sigma I$ has a bounded inverse, so the distribution
  $Tu_{\sigma} \in L^2(X_{sr} \times \bR)$. The inverse is bounded by
  $1/\sigma$, so the estimates \ref{eqn:snorm}, \ref{eqn:unorm},
  \ref{eqn:tcommst}, and \ref{eqn:tsscommnorm} combine to yield the
  bound \ref{eqn:tunorm}.

@@

\noindent {\bf Remark:} Note that the homogeneity in $\lambda$ of the
expressions in equations \ref{eqn:g0} and \ref{eqn:g1} results from
the proportionality between $\lambda$ and the regularization weight in
the definitions \ref{eqn:reg} and \ref{eqn:leading}. Otherwise, such
simple asymptotic relations as the conclusions
of Proposition \ref{thm:leading} would not obtain.

@@

\begin{prop}
  \label{thm:remest}
  Under the hypotheses of Theorem \ref{thm:main},
  \begin{eqnarray}
    \|\ul-\uzl\| &\le& C \lambda^{\frac{1}{2}}\|\dl\|\label{eqn:diffu} \\
    \|T(\ul-\uzl)\| &\le& C \lambda^{\frac{1}{2}}\|\dl\|\label{eqn:difftu}
  \end{eqnarray}
\end{prop}
